Hierarchical Phrase-Based Grammar Extraction in Joshua

### 2009-12-02 papers/mt_marathon10_SuffixArray/paper.tex
Version submitted to MT Marathon
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1209 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

\begin{abstract}
The abstract of the article...
\end{abstract}

@@

Significant recent research in statistical machine translation has focused on modeling translation based on contiguous strings of words, called {\textit phrases}, in the source language and corresponding phrases in the target language. Such phrase-based models \cite{Koehn2003,Marcu2002,Och2004,Och1999} have been very successful, and state-of-the-art machine translation systems are commonly based on these approaches.

@@

\newcite{Lopez2008} provides a fast implementation of SCFG grammar extraction for Hiero which uses suffix arrays. This implementation allows Hiero to use an aligned parallel corpus at run-time in lieu of a pre-extracted SCFG. However, this implementation is not available as free open source software. Cunei \cite{Phillips2009}\footnote{\url{http://sourceforge.net/projects/cunei}} is a statistical open-source EBMT system that uses suffix arrays to extract relevant phrase pairs from an aligned parallel corpus at run-time.

\section{Implementation: Data Structures and algorithms}
\label{data-structures}

@@

@@

Once the source and target corpus arrays are available, the corresponding {\bf suffix array}s can be constructed. Given a corpus array $c$ and a symbol table, a second array is created of equal size to the corpus array. This array $s$ is initialized such that $s[x] = x]$. Where each integer in $c$ represents a word string, each integer in $s$ represents an index into $c$. The contents of $s$ are sorted, using the element comparison function defined in figure \ref{sort}. After sorting, the indices of all instances in the corpus of any given phrase are located in a contiguous segment in the suffix array $s$.

@@

\begin{figure}
\begin{verbatimtab}[4]
<project name="Joshua Grammar Extraction in Ant" default="extract">

@@

</project>
\end{verbatimtab}
\caption{Sample Apache Ant build file defining how to extract a translation grammar from an aligned parallel corpus. This example assumes that the files {\tt source.txt}, {\tt target.txt}, {\tt align.txt} and {\tt target-test.txt} already exist in the directory specified in the {\tt experiment.dir} variable. Running ant with this file will create the directory {\tt model.josh} and extract {\tt test-set-specific.grammar}.}
\end{figure}

Grammar extraction using command line.

Grammar extraction using Apache Ant \cite{ant1.7.1}.

\begin{figure}
\begin{description}

@@

	\item[outputFile] Path where extracted grammar will be placed. This grammar will consist of all of the rules required to translate the sentences in the test file defined in the {\tt testFile} option.
\end{description}
\caption{Mandatory options must be configured in the {\tt extractRules} target.}
\end{figure}

\begin{figure}
\begin{description}

@@

	\item[cacheSize]

@@

\end{description}
\caption{Options can be configured in the {\tt extractRules} target.}
\end{figure}

\begin{figure}
\begin{description}

@@

	\item[requireTightSpans] Boolean option indicates whether

\end{description}
\caption{Options can be configured in the {\tt extractRules} target to make rule extraction behave like the Hiero suffix array rule extractor {\protect \cite{Lopez2008} } instead of the behaving according to the rule extraction originally defined in {\protect \newcite{Chiang2005}} .}
\end{figure}

\begin{figure}
\begin{description}
	\item[keepTree] Boolean option indicates whether the prefix tree should persist from sentence to sentence during grammar extraction. If set to false, a new prefix tree will be created to process each sentence in the test file. Default value is ``true".

@@

\end{description}
\caption{Additional options can be configured in the {\tt extractRules} target to affect behavior of the prefix tree.}
\end{figure}

\begin{figure}
\begin{description}
	\item[sourceFileName] Path to binary file containing integer-ized memory-mappable source language training corpus. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/source.corpus}.

	\item[targetFileName] Path to binary file containing integer-ized memory-mappable target language training corpus. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/target.corpus}.

	\item[alignmentsFileName] Path to binary file containing integer-ized memory-mappable source-to-target word alignments data. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/alignment.grids}. The format of this file is defined in {\tt joshua.corpus.alignment.AlignmentGrids\#writeExternal} and in the {\tt joshua.corpus.alignment.MemoryMappedAlignmentGrids} constructors.
\end{description}
\caption{Additional options can be configured in the {\tt extractRules} target. These options are normally set automatically as a result of setting the {\tt joshDir} option. They are listed here for completeness.}
\end{figure}

>>>>

\begin{abstract}
While example-based machine translation has long used corpus information at run-time, statistical phrase-based approaches typically include a pre-processing stage where an aligned parallel corpus is split into phrases, and parameter values are calculated for each phrase using simple relative frequency estimates. This paper describes an open source implementation of the crucial algorithms presented in \cite{Lopez2008} which allow direct run-time calculation of SCFG translation rules in Joshua.
\end{abstract}

@@

Significant recent research in statistical machine translation has focused on modeling translation based on contiguous strings of words, called {\it phrases}, in the source language and corresponding phrases in the target language. Such phrase-based models \cite{Koehn2003,Marcu2002,Och2004,Och1999} have been very successful, and state-of-the-art machine translation systems are commonly based on these approaches.

@@

\newcite{Lopez2008} provides a fast implementation of SCFG grammar extraction for Hiero which uses suffix arrays. This implementation allows Hiero to use an aligned parallel corpus at run-time in lieu of a pre-extracted SCFG. However, this implementation is not available as generally available open source software. Cunei \cite{Phillips2009}\footnote{\url{http://sourceforge.net/projects/cunei}} is a statistical open-source EBMT system that uses suffix arrays to extract relevant phrase pairs from an aligned parallel corpus at run-time.

\section{Implementation: Data Structures and Algorithms}
\label{data-structures}

@@

\ \\

@@

Once the source and target corpus arrays are available, the corresponding {\bf suffix array}s can be constructed. Given a corpus array $c$ and a symbol table, a second array is created of equal size to the corpus array. This array $s$ is initialized such that $s[x] = x$. Where each integer in $c$ represents a word string, each integer in $s$ represents an index into $c$. The contents of $s$ are sorted, using the element comparison function defined in figure \ref{sort}. After sorting, the indices of all instances in the corpus of any given phrase are located in a contiguous segment in the suffix array $s$.

@@

% Caching
Ideally, once translation rules have been extracted for a given source phrase, those rules would be stored and not calculated again. Unfortunately, memory constraints typically dictate that not all rules are stored. Rather than storing the translation rules for a given source phrase at the corresponding node in the prefix tree, a single least-recently-used (LRU) cache is maintained. This cache maps from source phrase to the corresponding set of translation rules.

% Memory mapping
Another technique used to save memory is the option of using memory-mapped data structures. Memory-mapped version of the corpus array, suffix array, and alignment grids data structures are implemented and used by default.

\begin{figure}[hptb]
\begin{verbatimtab}[4]
<project name="Joshua Grammar Extraction in Ant" default="extract">

@@

</project>
\end{verbatimtab}
\caption{Sample Apache Ant build file defining how to extract a translation grammar from an aligned parallel corpus. This example assumes that the files {\tt source.txt}, {\tt target.txt}, {\tt align.txt} and {\tt target-test.txt} already exist in the directory specified in the {\tt experiment.dir} variable. Running ant with this file will create the directory {\tt model.josh} and extract {\tt test-set-specific.grammar}.}
\label{ant-file}
\end{figure}

%\begin{figure}
%\begin{description}
%	\item[sourceFileName] Path to binary file containing integer-ized memory-mappable source language training corpus. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/source.corpus}.

%	\item[targetFileName] Path to binary file containing integer-ized memory-mappable target language training corpus. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/target.corpus}.

%	\item[alignmentsFileName] Path to binary file containing integer-ized memory-mappable source-to-target word alignments data. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/alignment.grids}. The format of this file is defined in {\tt joshua.corpus.alignment.AlignmentGrids\#writeExternal} and in the {\tt joshua.corpus.alignment.MemoryMappedAlignmentGrids} constructors.
%\end{description}
%\caption{Additional options can be configured in the {\tt extractRules} target. These options are normally set automatically as a result of setting the {\tt joshDir} option. They are listed here for completeness.}
%\end{figure}

Given a word-aligned parallel corpus, the first step in extracting a grammar, either at run-time or during a pre-processing stage, is to compile the memory-mappable data structures to binary files on disk. The {\tt joshua.corpus.suffix\_array.Compile} program takes four parameters: source corpus text file, target corpus text file, word alignments text file, and an output directory path. This program may be run directly, or may be accessed as an Apache Ant task (see figure \ref{ant-file}). The output directory, by convention, is named with the suffix .josh; this directory will contain binary representations of the symbol table, source and target corpus arrays, and source and target suffix arrays.

Once these binary files have been constructed into a .josh directory, the decoder can be configured to extract translation rules directly at run-time. To run the decoder in this mode, specify the .josh directory in the {\tt tm\_file=$\ldots$} line of the Joshua configuration file.

It is often useful (especially during MERT) to explicitly extract a test-set specific grammar in a pre-processing stage. The {\tt joshua.prefix\_tree.ExtractRules} program performs this task. When run directly, this program accepts either three arguments (a compiled .josh directory, file name for grammar to extract, and test file) or five arguments (source corpus text file, target corpus text file, word alignments text file, file name for grammar to extract, and test file). This program can also be accessed as an Apache Ant task (see figure \ref{ant-file}). The following subsections document the mandatory and optional parameters that can be passed to this program through the {\tt extractRules} ant task.

\subsection{Mandatory parameters}
%\begin{figure}
\begin{description}

	\item[outputFile] Path where extracted grammar will be placed. This grammar will consist of all of the rules required to translate the sentences in the test file defined in the {\tt testFile} option.

@@

\end{description}

The following parameters may be specified instead of the {\tt joshDir} parameter.

\begin{description}
	\item[sourceFileName] Path to text file containing source corpus. %this value is automatically set. Otherwise, this  to  Path to binary file containing integer-ized memory-mappable source language training corpus. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/source.corpus}.

	\item[targetFileName] Path to text file containing target corpus.

	%\item[targetFileName] Path to binary file containing integer-ized memory-mappable target language training corpus. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/target.corpus}.

	\item[alignmentsFileName]  Path to text file containing word alignment data.

	%\item[alignmentsFileName] Path to binary file containing integer-ized memory-mappable source-to-target word alignments data. When selected, the {\tt joshDir} option sets the value of this option to {\tt joshDir/alignment.grids}. The format of this file is defined in {\tt joshua.corpus.alignment.AlignmentGrids\#writeExternal} and in the {\tt joshua.corpus.alignment.MemoryMappedAlignmentGrids} constructors.

\end{description}
%\caption{Mandatory options must be configured in the {\tt extractRules} target.}
%\end{figure}

\subsection{Optional parameters}
%\begin{figure}
\begin{description}

@@

	\item[cacheSize] Maximum number of source phrases for which translation rules will be maintained in the least-recently-used (LRU) cache.

@@

\end{description}
%\caption{Options can be configured in the {\tt extractRules} target.}
%\end{figure}

The following parameters can be configured in the {\tt extractRules} target to make rule extraction behave like the Hiero suffix array rule extractor {\protect \cite{Lopez2008} } instead of the behaving according to the rule extraction originally defined in {\protect \newcite{Chiang2005}} .

%\begin{figure}
\begin{description}

@@

	\item[requireTightSpans] Boolean option; if true, follow the heuristic from \cite{Chiang2005}: where multiple initial phrase pairs contain the same set of alignment points, consider only the smallest when performing rule extraction. For compatibility with \newcite{Lopez2008}, set this parameter to false. Default value is ``true".

\end{description}
%\caption{Options can be configured in the {\tt extractRules} target to make rule extraction behave like the Hiero suffix array rule extractor {\protect \cite{Lopez2008} } instead of the behaving according to the rule extraction originally defined in {\protect \newcite{Chiang2005}} .}
%\end{figure}

Additional options can be configured in the {\tt extractRules} target to affect behavior of the prefix tree.

%\begin{figure}
\begin{description}
	\item[keepTree] Boolean option indicates whether the prefix tree should persist from sentence to sentence during grammar extraction. If set to false, a new prefix tree will be created to process each sentence in the test file. Default value is ``true".

@@

\end{description}
%\caption{Additional options can be configured in the {\tt extractRules} target to affect behavior of the prefix tree.}
%\end{figure}
